Uncovering underlying high-level musical content in the time domain \\ \small Leveraging self-supervised deep neural networks, inductive bias, and aural skills to learn deep audio embeddings with applications to MIR downstream tasks

### 2023-05-31 MasterThesis.tex
Updates from Overleaf

\usepackage[left=3cm,right=3cm,top=3cm,bottom=3cm]{geometry} %Margins
\usepackage{pdfpages}
\usepackage{hyperref}
\usepackage{listings}
\usepackage{xcolor}
\usepackage{setspace}
\usepackage{tocloft}
\usepackage{amsmath}
\usepackage{chngcntr }
\usepackage[toc,page]{appendix}
\usepackage[T1]{fontenc}
\usepackage[nottoc]{tocbibind}
\usepackage[compact]{titlesec}
\usepackage{float}

@@

\usepackage[utf8]{inputenc}
\usepackage{amsmath, amssymb, latexsym}
\usepackage{sidecap}

@@

%Title parameters
\title{%
  Uncovering underlying high-level musical content in the temporal domain \\
  \small Leveraging self-supervised deep neural networks, inductive bias, and aural skills to learn music embeddings with applications to MIR downstream tasks}

@@

\newpage
\input{appendix/appendix}

>>>>

\usepackage[left=3cm,right=3cm,top=3cm,bottom=3cm]{geometry} %Margins
\usepackage{pdfpages}
\usepackage{hyperref}
\usepackage{listings}
\usepackage{xcolor}
\usepackage{setspace}
\usepackage{tocloft}
\usepackage{amsmath}
\usepackage{chngcntr }
\usepackage[toc,page]{appendix}
\usepackage[T1]{fontenc}
\usepackage[nottoc]{tocbibind}
\usepackage[compact]{titlesec}
\usepackage{float}
\usepackage{booktabs}
\usepackage{makecell}

@@

\usepackage[utf8]{inputenc}
\usepackage{amsmath, amssymb, latexsym}
\usepackage{sidecap}
\usepackage{tabularx}

@@

%Title parameters
\title{%
  Uncovering underlying high-level musical content in the time domain \\
  \small Leveraging self-supervised deep neural networks, inductive bias, and aural skills to learn deep audio embeddings with applications to MIR downstream tasks}

@@

\newpage
\input{appendix/appendix}
\newpage

### 2023-05-31 abstract/abstract.tex
Updates from Overleaf

This thesis posits the existence of high-level musical concepts invariant to sonic qualities that evolve and unfolds through time. This idea parallels the nature of the symbolic domain, which maintains its essence despite being interpreted through diverse performances using various instruments and styles.

This research, conducted in collaboration with Epidemic Sound AB and the Music Technology Group (MTG) at Universitat Pompeu Fabra (UPF), employs self-supervised contrastive learning of musical representation to uncover the fundamental structure of Western tonal music.

Embracing a subtle novel approach, this process is facilitated by applying Siamese networks with a triplet loss, mimicking human aural skills in interpreting music to discern abstract and semantic musical elements, irrespective of the sonic qualities. The study replaces traditional acoustical features with deep audio embeddings to compute high-level, sound-agnostic, and content-sensitive music identification.

Results XXXXXX TBC

The preliminary results suggest that our approach to learning music-informed embeddings holds significant potential for nearly all MIR downstream tasks. Music-motivated embeddings represent a promising technique, adaptable potentially to other tasks hindered by data scarcity. This method could significantly influence advancements in intelligent music recommendation systems and the efficient enforcement of intellectual property rights.

>>>>

This thesis posits the existence of invariant high-level musical concepts that persist through changes in sonic qualities, akin to the symbolic domain where essence endures despite varying interpretations through different performances, instruments, and styles, among many other variables.

In collaboration with Epidemic Sound AB and the Music Technology Group (MTG) at Universitat Pompeu Fabra (UPF), we used self-supervised contrastive learning to uncover the underlying structure of Western tonal music. We applied deep convolutional neural networks and a triplet loss function, identifying abstract and semantic musical elements without relying on their sonic qualities. This way, we replaced traditional acoustic features with deep audio embeddings, paving the way for high-level, sound-agnostic, and content-sensitive music identification.

Our cognitively-based approach for learning embeddings focuses on using full resolution data and preserving high-level musical information. Drawing upon our domain expertise, we developed robust transformations to encode heuristic musical concepts that should remain constant. This novel approach aims to reconcile music and machine learning, enhancing machine listening models' efficacy. Preliminary results suggest that our musically-informed technique has significant potential for boundary detection tasks and possibly nearly all MIR downstream tasks that are not purely sonic-based. Consequently, music-motivated embeddings appear promising and potentially adaptable to other tasks constrained by data scarcity.

### 2023-05-31 figures/neural networks/manifold.tex
Updates from Overleaf

    \caption[High-level diagram of dimensionality reduction and latent space representation.]{Dimensionality reduction and latent space representation: Mapping between the original high-dimensional space ${\cal X}$ and the lower-dimensional latent space ${\cal F}$ using functions $f$ and $g$.}
    \label{fig:my_label}
\end{figure}

>>>>

    \caption[High-level diagram of dimensionality reduction and latent space representation \cite{tikz}.]{\small{Dimensionality reduction and latent space representation: Mapping between the original high-dimensional space ${\cal X}$ and the lower-dimensional latent space ${\cal F}$ using functions $f$ and $g$.}}
    \label{fig:manifold}
\end{figure}

### 2023-05-31 introduction/Subsections/Assumptions.tex
Updates from Overleaf

This thesis posits the existence of high-level musical concepts invariant to sonic qualities that evolve and unfolds through time, such as melody, harmony, rhythm, form, motif, counterpoint, etc. This idea parallels the nature of the symbolic domain (sheet music), which maintains its essence despite being interpreted through diverse performances using various instruments and styles. While a single waveform might have a limited number of sheet music representations that adequately capture its musical components, a single piece can be interpreted in infinite ways through different instruments, voices, tempos, and interpretations.

Abstract high-level features can be discerned regardless of waveform or production style. This approach provides an objective, comprehensive insight into a composition's musical content, much like analyzing sheet music uncovers its behavioral foundation.

For example, despite the differences in cultural contexts, stylistic elements, time signatures, key signatures, and sonic properties, we propose that figures \ref{fig:mahler} and \ref{fig:giant_steps} share a similarity. Both pieces carry the same musicological fingerprint, so to say, derived from the envelope of their melodic and harmonic contours through non-diatonic major thirds.

This concept finds grounding in the theories of Kurt Koffka, a founder of the Gestalt school of psychology, who proposed a holistic approach to understanding complex forms, contrary to the structuralist practice of reducing mental processes into essential elements \cite{Koffka2013PrinciplesPsychology}. When applied to music, the principles of Gestalt psychology offer insights into how we perceive, organize, and interpret musical patterns. They suggest that our minds perceive auditory input much like visual input, seeking patterns and structures. This understanding significantly contributes to the comprehension of cognitive processes in musical perception and organization, influencing music theory, cognition, and therapy \cite{Lerdahl1985AMusic}.

@@

\begin{figure}
    \centering
    \scalebox{0.9}{
    \includegraphics[width=\textwidth]{figures/images/train interlude score.png}
        }
    \caption[Last Train Home]{\small{Pat Metheny's Last Train Home interlude.}}
    \label{fig:last_train}
\end{figure}

>>>>

This thesis proposes the existence of high-level musical concepts, including melody, harmony, rhythm, form, motif, and counterpoint. These concepts, despite changes in sonic qualities, demonstrate enduring independence. Additionally, these elements are interactive, continually evolving within the musical context. This echoes the nature of the symbolic domain, such as sheet music, which retains its essence despite varied interpretations across diverse performances, instruments, and styles. While a single waveform might correspond to a finite set of sheet music representations encapsulating its musical components, one piece of music can be interpreted infinitely, varying by instruments, voices, tempos, and performance styles.

Abstract high-level features can be discerned irrespective of the waveform or production style, providing an objective, comprehensive insight into the musical content of a composition. This approach is akin to analyzing sheet music to uncover its foundational behaviors.

These features enable listeners to identify and recreate musical structures by ear, fostering deeper engagement with music. Heuristics, mental shortcuts, or rules of thumb facilitate rapid comprehension of complex musical concepts, bringing underlying patterns and structures into focus.

For example, despite differences in cultural contexts, stylistic elements, time signatures, key signatures, and sonic properties, we propose a similarity between figures \ref{fig:mahler} and \ref{fig:giant_steps}. Both pieces exhibit the same musicological fingerprint, so to speak, derived from the envelope of their melodic and harmonic contours through non-diatonic major thirds.

This concept aligns with the theories of Kurt Koffka, a founding member of the Gestalt school of psychology, who advocated for a holistic approach to understanding complex forms, as opposed to the structuralist practice of dissecting mental processes into their elemental components \cite{Koffka2013PrinciplesPsychology}. In the context of music, Gestalt psychology principles suggest our minds process auditory input similarly to visual input, by seeking patterns and structures. This understanding significantly enhances the comprehension of cognitive processes in musical perception and organization, influencing music theory, cognition, and therapy \cite{Lerdahl1985AMusic}.

@@

### 2023-05-31 introduction/Subsections/Introduction.tex
Updates from Overleaf

Music \footnote{In this thesis, "music" refers to the Western tradition, characterized by conventions, practices, and aesthetics that have developed primarily in Europe and North America. This usage is not intended to deny the diversity and richness of other musical traditions worldwide but instead reflects the present study's focus on a specific cultural context.}, an essential component of human culture, history, and society; it has existed since ancient times. It reflects the changes in our cultures, traditions, and beliefs as it evolves. Studying music, therefore, has always been a challenging and complex task due to its subjective nature, relative and mutable ground truth, and the sheer variety of musical styles, genres, and worldwide cultures.

For millennia, people have sought to understand the underlying structure of music. Music has captivated and intrigued us from the ancient Greeks to the present day. This rich music theory and analysis history form a valuable foundation for guiding and evaluating Music Information Retrieval (MIR) algorithms and models.

MIR has emerged over the last few decades as an interdisciplinary field that combines engineering, musicology, and neuroscience. It aims to develop algorithms and techniques for automatically analyzing, classifying, and retrieving musical information, primarily from audio signals. While MIR research has traditionally focused on technical aspects such as Digital Signal Processing (DSP) and pattern recognition, it may have overlooked other relevant domains, including perception, cognition, music theory, aural skills, musicology, and their subfields.

>>>>

Music is essential to human culture, history, and society; it has existed since ancient times. It reflects the changes in our cultures, traditions, and beliefs as they evolve \footnote{In this thesis, "music" refers to the Western tradition, characterized by conventions, practices, and aesthetics that have developed primarily in Europe and North America. This usage does not deny the diversity and richness of other musical traditions worldwide but reflects the present study's focus on a specific cultural context.}. Studying music, therefore, has always been a challenging and complex task due to its subjective nature, relative and mutable ground truth, and the sheer variety of musical styles, genres, and worldwide cultures.

For millennia, people have sought to understand the underlying structure of music. Music has captivated and intrigued us from the ancient Greeks to the present day. This rich music theory and analysis history form a valuable foundation for guiding and evaluating MIR algorithms and models.

MIR has emerged over the last few decades as an interdisciplinary field that combines engineering, musicology, and neuroscience. It aims to develop algorithms and techniques for automatically analyzing, classifying, and retrieving musical information, primarily from audio signals.

### 2023-05-31 introduction/Subsections/Motivation.tex
Updates from Overleaf

As a music enthusiast, professional musician, and scholar, I am passionately committed to exploring the foundational elements of musical composition, investigating the diverse techniques employed in creating musical works, and deciphering the intricate relationships among them.

This passion fuels my quest for knowledge on extracting valuable information embedded within music in any domain, being my main interest in the audio and symbolic domains, contributing to developing more sophisticated and effective music-related technologies and solutions in the industrial landscape.

\subsubsection{Academic and industrial motivation}

@@

The emergence of self-supervised models that learn embedding spaces for retrieving musical content from audio signals has unveiled new opportunities in academic research. By capturing high-level semantic information from raw audio data in an unsupervised manner, these models can potentially revolutionize various aspects of the music industry.

@@

Purely mathematical approaches may not always reflect how humans perceive music and may not capture the subtle variations in timing or harmonic relationships critical to musical perception. We argue that MIR research needs to incorporate a more balanced approach that considers the interdisciplinary nature of music and the importance of other domains beyond digital signal processing.

I have consistently been inspired by musicians and researchers who strive to uncover musical ground truth within an existing tonal paradigm, such as Heinrich Schenker and his school of thought \cite{Komar1959SchenkersStructure}, and particularly by challenging it, exemplified by Arnold Schoenberg \cite{Samson1974SchoenbergsMusic}, George Russell \cite{LydianRussell} or Ernst Levy \cite{LevyAHarmony}. These approaches aim to identify abstract concepts that reinforce or disrupt the tonal foundation. Their ultimate goal is to advance the tonal landscape, providing musicians with a dependable playground for growth, development, and further understanding tonal and atonal paradigms.

>>>>

As a music enthusiast, musician, and scholar, I am passionately committed to exploring the foundational elements of musical composition, investigating the diverse techniques employed in creating musical works, and deciphering the intricate relationships among them.

This passion fuels my quest for knowledge on extracting valuable information embedded within music in any domain, being my main interest in the audio and symbolic domains \cite{Crestel2018AOrchestration}\cite{WeilAutomaticSignals}, contributing to developing more sophisticated and effective music-related technologies and solutions in the industrial landscape.

\subsubsection{Industrial motivation}

@@

The emergence of self-supervised models that learn embedding spaces for retrieving musical content from audio signals has unveiled new opportunities. By capturing high-level semantic information from raw audio data in an unsupervised manner, these models can potentially revolutionize various aspects of the music industry.

@@

While MIR research has traditionally focused on technical aspects, it may have overlooked other relevant domains, including perception, cognition, music theory, aural skills, musicology, and their subfields. Purely mathematical approaches may not always reflect how humans perceive music and may not capture the subtle variations in timing or harmonic relationships critical to musical perception. We argue that MIR research needs to incorporate a more balanced approach that considers the interdisciplinary nature of music and the importance of other domains beyond DSP.

I have consistently been inspired by musicians and researchers who strive to bridge the gap and uncover musical ground truth within an existing tonal paradigm, such as Heinrich Schenker and his school of thought \cite{Komar1959SchenkersStructure}, and particularly by challenging it, exemplified by Arnold Schoenberg \cite{Samson1974SchoenbergsMusic}, George Russell \cite{LydianRussell} or Ernst Levy \cite{LevyAHarmony}. These approaches aim to identify abstract concepts that reinforce or disrupt the tonal foundation. Their ultimate goal is to advance the tonal landscape, providing musicians with a dependable playground for growth, development, and further understanding of tonal and atonal paradigms.

### 2023-05-31 introduction/Subsections/Previous work.tex
Updates from Overleaf

These aspects include the utilization of multi-level music segmentation through deep audio embeddings to replace manually engineered features \cite{SalamonDeepSegmentation}, the enhancement of environmental sound classification \cite{Kim2020OneStrategies}, the improvement of vocal-centric music tasks using cross-domain audio embedding \cite{Kim2021LearningLoss}, the enhancement of audio classification through a combination of task-specific and pre-trained features \cite{Hung2022Feature-informedClassification}, the creation of a music similarity search engine specifically for video producers \cite{epidemic}, the improved performance in Music Emotion Recognition (MER) tasks \cite{KohComparisonRecognition}, embedding songs for similarity comparison based on artist identification \cite{contentmusicsimtriplet2020}, the resolution of the cross-modal text-to-music retrieval problem \cite{WonEmotionStories}, and music rearrangement \cite{Stoller2018IntuitiveTransitions}. This last innovative approach presents a promising avenue for automated music rearrangement, enhancing the process's efficiency and quality \cite{Plachouras2023MusicSegmentation}.

>>>>

These aspects include the utilization of multi-level music segmentation through deep audio embeddings to replace manually engineered features \cite{SalamonDeepSegmentation}, the enhancement of environmental sound classification \cite{Kim2020OneStrategies}, the improvement of vocal-centric music tasks using cross-domain audio embedding \cite{Kim2021LearningLoss}, the enhancement of audio classification through a combination of task-specific and pre-trained features \cite{Hung2022Feature-informedClassification}, the creation of a music similarity search engine specifically for video producers \cite{epidemic}, the improved performance in Music Emotion Recognition (MER) tasks \cite{KohComparisonRecognition}, embedding songs for similarity comparison based on artist identification \cite{contentmusicsimtriplet2020}, the resolution of the cross-modal text-to-music retrieval problem \cite{WonEmotionStories}, and automated music rearrangement \cite{Stoller2018IntuitiveTransitions}\cite{Plachouras2023MusicSegmentation}.

### 2023-05-31 methods/subsections/deep learning.tex
Updates from Overleaf

They leverage the central algorithm, backpropagation, which calculates gradients of the loss function related to network parameters through forward and backward pass processes, thereby optimizing model performance.

@@

Deep Neural Networks (DNNs), computational models rooted in biological neural systems, utilize layers of artificial neurons for tasks such as pattern recognition, classification, and regression. DNNs evolve by adjusting connections between neurons during learning, a concept initiated in the mid-20th century but modernized with Frank Rosenblatt's single-layer perceptron in 1957, capable of learning linearly separable patterns.

@@

Another crucial component in the operation of DNNs is the activation function. This function is applied to a neuron's inputs, transforming the input's summed product and weights into an output passed onto the next layer. It introduces non-linearity into the network, enabling it to learn complex patterns and relationships from the data.

@@

%Colour CNN
\input{figures/neural networks/colour CNN}

>>>>

They leverage the central algorithm, backpropagation, which calculates gradients of the loss function related to network parameters through forward and backward pass processes, thereby optimizing model performance \cite{Rumelhart1986LearningInternal}.

@@

Deep Neural Networks (DNNs), computational models rooted in biological neural systems, utilize layers of artificial neurons for tasks such as pattern recognition, classification, and regression. DNNs evolve by adjusting connections between neurons during learning, a concept initiated in the mid-20th century but modernized with Frank Rosenblatt's single-layer perceptron in 1957, capable of learning linearly separable patterns \cite{Rosenblatt1958ThePerceptron}.

@@

Another crucial component in the operation of DNNs is the activation function. This function is applied to a neuron's inputs, transforming the input's summed product and weights into an output passed onto the next layer. It introduces non-linearity into the network, enabling it to learn complex patterns and relationships from the data \cite{Goodfellow2016DeepLearning}.

@@

### 2023-05-31 methods/subsections/implementation.tex
Updates from Overleaf

\section{Implementation details}
\subsection{Contrastive Learning of Musical Representations}

@@

For this task, a Triple Siamese Network is suggested, a model architecture proved efficient for music similarity retrieval tasks \cite{contentmusicsimtriplet2020}, intending to minimize the loss function between an anchor, positive, and negative sample, achieved through online triplet mining \cite{Sikaroudi2020OfflinePatches}. The overarching aim is to train a model in a self-supervised setting, which can effectively distinguish between similar and dissimilar samples while considering their underlying composition and production texture.

@@

The SampleCNN model \cite{Lee2018SampleCNN:Classification} takes 1D input data with a single channel and follows the next architecture:

@@

\begin{equation}
y_i = \frac{1}{N} \sum_{j=1}^{N} x_{ij}
\end{equation}

Here, $y_i$ is the output tensor after the average pooling process. Each element $x_{ij}$ is part of the input tensor. $N$ is the count of elements in the second dimension of the input tensor, while $i$ and $j$ are used to iterate over the first and second dimensions of the tensors, respectively.

@@

\subsubsection{Optimizer}

@@

We believe our experiments can benefit from time-domain processing because it allows more natural handling of temporal patterns and sequences in the data while avoiding windowing artifacts.

Mel-spectrogram-based approaches have proved useful in numerous audio-related machine-learning tasks using CNN, as their frequency-domain summarization is efficient and not easily learnable through data-driven approaches \cite{Kim2020OneStrategies}. That being said,  they exhibit limitations concerning their representation of perceptual similarity: the spectrogram of an audio signal, denoted as $\phi(x)$ for audio signal $x$ and $\phi(y)$ for audio signal $y$, captures the distribution of frequencies as they vary over time. A small difference between $\phi(x)$ and $\phi(y)$ generally signifies high perceptual similarity from a human listener's perspective. In other words, if two audio signals have similar spectrograms, they are likely to sound alike when heard.

However, the inverse is not consistently true. There exist instances where two audio signals can have significantly different spectrograms, i.e., a high spectrogram distance between $\phi(x)$ and $\phi(y)$, but are still perceived as musically similar by human listeners. This discrepancy is due to the complexity of human auditory perception that goes beyond just the distribution of frequencies over time \cite{Mesostructures2023}.

Our perception of sound involves several other factors, including temporal patterns, phase relationships between frequencies, and even higher-level musical structures. Therefore, even if two sounds have very different spectrograms, they might still be perceived as similar if they share these other characteristics. This underlines the inherent limitation of using only spectrogram distance to measure high-level music similarity, as it may not always produce results consistent with human perceptual judgments.

@@

Let $g \in [-12, 0]$ represent the gain, $\Delta p \in [-1200, 1200]$ the pitch shift, $h_R[n]$ the impulse response of the reverb with parameters in $[0, 100]$, $h_C[n]$ the impulse response of the chorus with parameters determined by the specified ranges, $d \in [0, 30]$ the overdrive parameter, $\alpha \in [0.9, 1.1]$ the speed change factor, $\beta \in [0.9, 1.1]$ the stretch factor, $f_m \in [0.1, 100]$ the modulation frequency of the tremolo, and $d_m \in [1, 101]$ the depth of the tremolo.

These effect selections behave and affect the signal in several domains, adding aggressive transformations while preserving high-level musical content.

\begin{itemize}
\item \textbf{Amplitude effects:} Amplitude or gain modification.
\begin{itemize}
    \item Gain: Adjusts the overall amplitude of the signal by a constant factor $g \in [-12, 0]$ dB.
\end{itemize}

\item \textbf{Time-domain effects:} Alters the signal's timing or duration.
\begin{itemize}
    \item Speed change: Alters the playback speed of the signal by a factor $\alpha \in [0.9, 1.1]$.
    \item Stretch: Changes the signal duration without affecting its pitch by a factor $\beta \in [0.9, 1.1]$.
\end{itemize}

\item \textbf{Frequency-domain effects:} Alters the frequency content or pitch of the signal.
\begin{itemize}
    \item Pitch shift: Changes the signal's pitch by $\Delta p \in [-1200, 1200]$ cents. Two-octave range, one octave higher to one octave lower.
\end{itemize}

\item \textbf{Nonlinear effects:} Harmonic distortion or other complex changes.
\begin{itemize}
    \item Overdrive: Adds harmonic distortion to the signal based on the parameter $d \in [0, 30]$.
\end{itemize}

\item \textbf{Modulation effects:} Modulation using a low-frequency oscillator or control signal.
\begin{itemize}
    \item Chorus: Applies a varying time delay to the signal, resulting in a richer, thicker sound. The specified ranges determine the chorus parameters.
    \item Tremolo: Modulates the signal's amplitude at a modulation frequency $f_m \in [0.1, 100]$ Hz and depth $d_m \in [1, 101]$.
\end{itemize}

\item \textbf{Reverberation effects:} Acoustic reflections simulation and reverberations of physical space.
\begin{itemize}
    \item Reverb: Applies an impulse response $h_R[n]$ that simulates the reverberation of space with parameters in the range $[0, 100]$.
\end{itemize}

\item \textbf{Noise effects:}
\begin{itemize}
    \item Additive noise: Adds a noise signal $N[n]$ with a specific signal-to-noise ratio (SNR) in the range $[12, 100]$.
\end{itemize}
\end{itemize}

The positive signal $P[n]$ is generated as follows:

@@

where $\ast$ denotes convolution, $h_{G}[n] = g \delta[n]$, $h_{D}[n]$ is the impulse response of the overdrive effect, $h_{P_t}[n]$ is the impulse response of the pitch shift effect, $h_{S}[n]$ represents the impulse response of the speed change effect, $h_{T}[n]$ is the impulse response of the stretch effect, and $h_{T_m}[n] = (1 - d_m \cos(2 \pi f_m n))\delta[n]$ is the impulse response of the tremolo effect. The noise signal $N[n]$ is added with a specific signal-to-noise ratio (SNR) in the range $[12, 100]$ in decibels.

@@

The triplet loss function used in Siamese networks guides the learning process. It seeks to minimize the distance between the anchor and positive instances while maximizing the distance between the anchor and the negative instances. A margin parameter is included in the loss function to ensure a minimum separation between positive and negative instances in the embedding space.

@@

The goal of minimizing this loss function is to learn discriminative embeddings, where similar examples are grouped closely together and dissimilar examples are placed farther apart in the embedding space.

>>>>

\section{Implementation details}

Our approach could be broadly categorized as cognitive modeling, as it strives to mimic human cognitive processes and problem-solving through a computerized model. We champion exposure-based learning in music and active engagement with an array of musical styles, genres, techniques, and learning methods to foster well-rounded musical proficiency and efficient performance with novel data in practical applications.

\subsection{Contrastive Learning of Musical Representations}

@@

For this task, a Triple Siamese Network is suggested, a model architecture proved efficient for music similarity retrieval tasks \cite{contentmusicsimtriplet2020}, intending to minimize the loss function between an anchor, positive, and negative sample, achieved through online triplet mining \cite{Sikaroudi2020OfflinePatches}. The overarching aim is to train a model in a self-supervised setting, which can effectively distinguish between similar and dissimilar samples.

@@

The SampleCNN model \cite{Lee2018SampleCNN:Classification} is a CNN designed for raw waveform audio data, treating each audio sample as an independent channel and applying 1-dimensional convolution along the temporal axis. Its implementation is an adaptation of \cite{CLMR2021} using PyTorch \cite{Paszke2019PyTorch:Library} and PyTorch Lightning \cite{PyTorchDocumentation}.

This fully convolutional model reduces computational requirements and learns features at different scales through its multi-resolution architecture \ref{tab:samplecnn}.

@@

@@

\subsection{Optimizer and learning rate}

In our research, we've employed the AdamW optimizer \cite{Loshchilov2017DecoupledRegularization}, an optimized variant of the widely-used Adam optimizer for training neural networks. AdamW adeptly balances the learning rate across network weights, providing an efficient strategy for weight decay management by isolating it from gradient updates.

\subsubsection{Learning rate}

The learning rate, set at 0.003, is a pivotal parameter dictating the step size at each iteration towards loss function minimization. It's a delicate balancing act— a high rate promises swift convergence with a risk of minimum overshoot, while a lower rate provides careful convergence but necessitates more iterations.

@@

Our experiments can benefit from incorporating time-domain processing, which allows for more natural handling of temporal patterns and sequences in the data while avoiding windowing artifacts. While mel-spectrogram-based approaches have proven useful in various audio-related machine-learning tasks, their limitations lie in their representation of perceptual similarity. Spectrograms capture the distribution of frequencies over time, and small differences in spectrograms generally indicate high perceptual similarity from a human listener's perspective. However, there are cases where two audio signals can have significantly different spectrograms but still be perceived as musically similar. This discrepancy is due to the complexity of human auditory perception, which encompasses temporal patterns, phase relationships between frequencies, and higher-level musical structures. Therefore, relying solely on spectrogram distance to measure high-level music similarity may not consistently align with human perceptual judgments \cite{Kim2020OneStrategies}\cite{Mesostructures2023}.

@@

In this analysis, various parameters represent different effects on the signal. We use $g \in [-12, 0]$ for gain, $\Delta p \in [-1200, 1200]$ for pitch shift, $h_R[n]$ for the reverb impulse response, $h_C[n]$ for the chorus impulse response, $d \in [0, 30]$ for overdrive, $\alpha \in [0.9, 1.1]$ for speed change, $\beta \in [0.9, 1.1]$ for stretch factor, $f_m \in [0.1, 100]$ for tremolo's modulation frequency, and $d_m \in [1, 101]$ for tremolo's depth. Each effect is applied aggressively but does not disrupt the high-level musical content.

\textbf{Amplitude effects}, like gain, modify the signal's overall amplitude with a constant factor $g$.

\textbf{Time-domain effects} such as speed change and stretch alter the signal's playback speed and duration without changing its pitch, respectively. The speed and stretch factors are $\alpha$ and $\beta$.

\textbf{Frequency-domain effects} adjust the frequency content or pitch of the signal, as in pitch shifting, which alters the pitch by $\Delta p$.

\textbf{Nonlinear effects} such as overdrive add harmonic distortion using the parameter $d$.

\textbf{Modulation effects}, including chorus and tremolo, use a control signal or low-frequency oscillator. The specified ranges determine the chorus parameters, and the tremolo's amplitude modulation frequency and depth are $f_m$ and $d_m$.

\textbf{Reverberation effects} simulate physical space's acoustic reflections and reverberations, like reverb that applies an impulse response $h_R[n]$. Finally, \textbf{Noise effects} add a noise signal $N[n]$ with a signal-to-noise ratio (SNR) in the range $[12, 100]$.

The positive signal $P[n]$ is generated according to the equation given below, where the symbols denote specific effects as indicated:

@@

The $\ast$ symbol represents convolution, and the various $h$ symbols indicate the impulse responses of specific effects. The noise signal $N[n]$ is included with a specific SNR.

@@

The triplet loss function used in Siamese-like networks guides the learning process. It seeks to minimize the distance between the anchor and positive instances while maximizing the distance between the anchor and the negative instances. A margin parameter is included in the loss function to ensure a minimum separation between positive and negative instances in the embedding space.

@@

As previously stated, the goal of minimizing this loss function is to learn discriminative embeddings, where similar examples are grouped closely together while dissimilar examples are placed farther apart in the embedding space.

### 2023-05-31 preface/preface.tex
Updates from Overleaf

%\addcontentsline{toc}{chapter}{Acknowledgement}
\begin{preface}
\pagenumbering{gobble}% Remove page numbers (and reset to 1)
(Optional, if used placed on a right page next to an empty left page)

I would like to dedicate this work to...
\newpage

\end{preface}

>>>>

%\addcontentsline{toc}{chapter}{Acknowledgement}
\begin{preface}
\pagenumbering{gobble}% Remove page numbers (and reset to 1)

Music and science have always been my two greatest passions.

When the opportunity to work at Epidemic Sound presented itself, it was as if the universe had conspired to bring my interests together, offering me the chance to explore the fascinating field of Music Information Retrieval (MIR) from an industrial perspective.

With the support and encouragement of my loved ones, I took a leap of faith and moved from Barcelona to Stockholm to embark on this exciting journey.

This work humbly aims to help untangle the complexities of music and offer a more musically-driven perspective to MIR. We aim to contribute to a deeper understanding of the intricate relationships within musical data by bridging the gap between music theory and computational analysis.

I am deeply grateful to the Epidemic Sound team for their guidance, expertise, and camaraderie throughout this project. I would also like to extend my heartfelt appreciation to my family for their unwavering support and belief in my abilities.

My motivation for writing this piece stems from a desire to grow as a musician, developer, and individual. I believe that by delving into the world of MIR, I can expand my horizons while making a meaningful contribution to the field.

This work is intended for any curious human being, regardless of their background in music or computer science. I hope it will inspire others to explore the fascinating intersection of these two disciplines.

The scope of this work encompasses a range of MIR tasks and methodologies, as well as discussions on the challenges and opportunities that arise within the field. We hope this work will serve as a starting point for further exploration and inspire new ideas in the realm of MIR.

\newpage
\end{preface}
